Baseline centering problem

### 2018-12-19 paper/tangotd.tex

The reason for this form is the following: $E$ is supposed to represent
distributions or functions over states. $f$ is supposed to map $s_2$ to
something like
$\1_{s_2}$ (up to a factor $1/\rho$ to compensate for sampling
probabilities). 
In
particular, $\E_{s_2\sim \rho}f(s_2)\psi(s_2)$ is supposed to represent
the vector with components $\psi(s_2)$, namely, the function $\psi$. $\green$ is supposed to map each
$s_1$ to the
the occupation
distribution of the random walk starting from $s_1$, namely,
$\transp{\1_{s_1}}\sum_t P^t$, which is $\transp{\1_{s_1}}\,\Delta^{-1}$
(as usual I
omit constants\footnote{To deal with constants, we probably have to be
able to predict \emph{differences} $\psi(s_2)-\psi(s_2')$, or just use
relative value functions and remove $\psi(s_0)$ for some $s_0$.}). Then the scalar product in
$E$ represents integration/summation over the state space.

@@

Intuitively, the middle Bellman equation does not create new starting
points or endpoints on paths, only works in the middle. (This corresponds
to the fixed image and kernel.) Meanwhile the left Bellman equation
creates starting points and the right Bellman equation creates endpoints
when observing a new transition $s\to s'$. This introduces spatial
smoothness over the image and kernel.

@@

>>>>

The reason for this form is the following: $E$ is supposed to represent
distributions or functions over states. $f$ is supposed to map $s_2$ to
something like
$\1_{s_2}$ (up to a factor $1/\rho$ to compensate for sampling
probabilities). 
In
particular, $\E_{s_2\sim \rho}f(s_2)\psi(s_2)$ is supposed to represent
the vector with components $\psi(s_2)$, namely, the function $\psi$. $\green$ is supposed to map each
$s_1$ to the
the occupation
distribution of the random walk starting from $s_1$, namely,
$\transp{\1_{s_1}}\sum_t P^t$, which is $\transp{\1_{s_1}}\,\Delta^{-1}$
(as usual I
omit constants\footnote{To deal with constants, we probably have to be
able to predict \emph{differences} $\psi(s_2)-\psi(s_2')$, or just use
relative value functions and remove $\psi(s_0)$ for some $s_0$.}). Then the scalar product in
$E$ represents integration/summation over the state space.
This representation keeps linearity of the map $R\mapsto V$, while
allowing for a nonlinear $f$ and $\green$.

@@

Intuitively, the middle Bellman equation does not create new starting
points or endpoints on paths, only works in the middle. (This corresponds
to the fixed image and kernel.) [TODO: though if I start with $Z=\Id$ (a
path from every state to itself) and apply Newton then I do create paths,
so the meaning of this is unclear and initialization might be important.
It might be important to initialize $F$ and $\green$ so that $Z\approx \Id$.] Meanwhile, the left Bellman equation
creates starting points and the right Bellman equation creates endpoints
when observing a new transition $s\to s'$. This introduces spatial
smoothness over the image and kernel.

@@

\paragraph{Variant.}
I think
everything works equally well if we just learn a kernel $Z(s_1,s_2)$ such
that
\begin{equation}
\transp{\1_{s_1}}\Delta^{-1}\psi\approx \E_{s_2\sim \rho} Z(s_1,s_2)
\psi(s_2)
\end{equation}
and this is still linear in $\psi$. This represents the inverse Laplace
operator by
\begin{equation}
Z=\sum_{s_1,s_2}\rho(s_2)  Z(s_1,s_2)\1_{s_1}\transp{\1_{s_2}}
\end{equation}

The left Bellman equation is
\begin{gather}
\transp{\1_{s_1}}\,\Delta^{-1}\psi=\transp{\1_{s_1}}\,\psi+\transp{\1_{s_1}}\,P\Delta^{-1}\psi
\qquad \Longrightarrow
\\
\E_{s_2\sim \rho}
\,Z(s_1,s_2)\psi(s_2)
\approx \psi(s_1)+ \E_{s'\sim P(s'|s_1)}
\E_{s_2\sim \rho}\,
Z(s',s_2)\psi(s_2)
\end{gather}
for any $s_1$ and $\psi$.
The right Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\,\Delta^{-1}\psi=\transp{\1_{s_1}}\,\psi+\transp{\1_{s_1}}\,\Delta^{-1}P\psi
\qquad \Longrightarrow
\\
\E_{s_2\sim \rho}
Z(s_1,s_2)\psi(s_2)
\approx \psi(s_1)+ \E_{s_2\sim \rho}\E_{s'_2\sim P(s'_2|s_2)}
Z(s_1,s_2)\psi(s'_2)
\end{gather}
These equations impose a kind of smoothness on $Z$ with respect to the
random walk $P$.

The Newton update is
\begin{equation}
Z\gets (1+\eta)Z-\eta Z^2+\eta \gamma ZPZ
\end{equation}
which after applying to $\1_{s_1}$ on the left yields
\begin{multline}
Z(s_1,s_2)\gets (1+\eta)Z(s_1,s_2)-\eta \E_{s_3\sim \rho}
Z(s_1,s_3)Z(s_3,s_2)\,+\\\eta \gamma \E_{s_3\sim\rho, s'_3\sim P(s'_3|s_3)}
Z(s_1,s_3)Z(s'_3,s_2)
\end{multline}
and we should find a way to initialize $Z$ so that $Z\approx \Id$ at time
$0$.
